Inefficient Concessions and Mediation

### 2023-08-06 draft_narrow.tex
Broaden conclusion language slightly
Add "public goods" in a few places where we only talk about peace

Examples of a \emph{no-concessions separating equilibrium} are most likely to be those where countries have traditionally had good relations but something has recently come in the way. Certain relationships between the USA and countries alienated by WikiLeaks -- such as Italy -- would be an example. Italy essentially laughed off the criticisms and reassured the USA of its commitment to their relationship. Warsaw Pact countries generally allowed the USSR similar leeway; it was better to have occasionally imperfect cooperation than conflict. Some events can disrupt trust without completely derailing cooperation.

@@

While parties generally are better off when concessions are efficient, we have shown that inefficiency may be useful. In particular, if concessions may be used to harm the giver materially, it may be better to give inefficient concessions. This explanation captures the hesitancy of nations to fully engage in moving forward with a peace process when there is mutual distrust.

@@

This model focuses on manipulative mediation, which is indeed used in resolving conflicts. However, the usefulness of mediation in this case relies on the mediator being both perfectly trustworthy and perfectly able to enforce concessions. Such an all-powerful mediator is a strong assumption, and while useful as a benchmark, it is unlikely to exist in all scenarios. It would be useful as a further exercise to examine the effectiveness of mediators with different strengths of enforcement capabilities.

@@

This paper shows that inefficient concessions may be used to achieve peace when efficient concessions are not sufficient to do so; it also demonstrates how a mediator can achieve peace in these situations when the parties to the conflict cannot do so on their own. Opportunities for further research clearly include looking at the roles of different types of mediation. Because of the many humanitarian, security and diplomatic implications of this work, both deeper and broader analysis is required.

>>>>

%Examples of a \emph{no-concessions separating equilibrium} are most likely to be those where countries have traditionally had good relations but something has recently come in the way. Certain relationships between the USA and countries alienated by WikiLeaks -- such as Italy -- would be an example. Italy essentially laughed off the criticisms and reassured the USA of its commitment to their relationship. Warsaw Pact countries generally allowed the USSR similar leeway; it was better to have occasionally imperfect cooperation than conflict. Some events can disrupt trust without completely derailing cooperation.

@@

While parties generally are better off when concessions are efficient, we have shown that inefficiency may be useful. In particular, if concessions may be used to harm the giver materially, it may be better to give inefficient concessions. This explanation captures the hesitancy of nations to fully engage in moving forward with a peace process or other cooperation when there is mutual distrust.

@@

This model focuses on manipulative mediation, which is indeed used in resolving conflicts and achieving cooperation. However, the usefulness of mediation in this case relies on the mediator being both perfectly trustworthy and perfectly able to enforce concessions. Such an all-powerful mediator is a strong assumption, and while useful as a benchmark, it is unlikely to exist in all scenarios. It would be useful as a further exercise to examine the effectiveness of mediators with different strengths of enforcement capabilities.

@@

This paper shows that inefficient concessions may be used to achieve peace or cooperation to provide of other public goods when efficient concessions are not sufficient to do so; it also demonstrates how a mediator can achieve cooperation in these situations when the parties to the conflict cannot do so on their own. Opportunities for further research clearly include looking at the roles of different types of mediation. Because of the many humanitarian, security and diplomatic implications of this work, both deeper and broader analysis is required.
